---
title: "Data Driven Emotion Dynamics Code"
author: "Monica Kullar"
output: html_document
---
Table of Contents
1. Pre-processing ESM data
2. Confirmatory groups GIMME analysis - clinical diagnostic subgroups
3. Exploratory groups GIMME analysis - data-driven subgroups
4. Final GIMME solution evaluation

```{r}
#packages used
library(dplyr)
library(data.table)
library(imputeTS)
library(pracma)

##GIMME packages
devtools::install_github("GatesLab/gimme/tree/master/gimme", force= TRUE, dependencies = TRUE)
#devtools::install_github("GatesLab/gimme/gimme", force=TRUE)
library(gimme)
library(perturbR)
```

#1. Pre-processing ESM data
```{r}
# Impute data in order to use standardized residuals of time-series data
esmdata <- read.csv("~/esmdata.csv") #read in esmdata from your local filepath

b1var <- esmdata
    shortem <- c("moniker", "time", "Happy_e","Enthusiastic","Pleased", "Relaxed", "Nervous", "Sad", "Irritated", "Angry", "Stressed", "MWoccur", "EmotionChronometry")
    b1var <- b1var[shortem]
    imputevar <- c("Happy_e","Enthusiastic","Pleased", "Relaxed", "Nervous", "Sad", "Irritated", "Angry", "Stressed", "MWoccur", "EmotionChronometry")
    b1var$moniker <- as.factor(b1var$moniker)
#loop through participants
for(i in levels(b1var$moniker)) {
    b1varimp <- b1var[b1var$moniker == i,]
        for(j in imputevar) {
          b1varimp[,j] <- na_ma(b1varimp[,j], k=2, weighting = "simple")
        }
    #save all files in folder for use at individual level
    savefile <- paste0 ("~/step1/", i, ".csv", sep = "") #save first preprocessed step here to a folder you name 'step1' (or change the name to what you would like)
    write.csv(b1varimp, file = savefile, row.names = FALSE)
    print(paste("Dataframe Saved:", i))
}
    # Reduce down the most highly correlating emotion variables used in analysis of input variables
    data <- rbindlist(lapply(list.files("~/step1/", full.names = TRUE), fread), fill = TRUE) #read in individual files from prior step
        data$moniker <- as.factor(data$moniker)
    #reduce down the most highly correlating emotion variables
    data$HighCorrNegative <- (data$Angry + data$Irritated)/2 #referred to as "Angry" in manuscript for ease of interpreting
    data$HighCorrPositive <- (data$Happy_e + data$Pleased)/2 #referred to as "Happy" in manuscript for ease of interpreting
    data <- as.data.frame(data)
    corrred <- c("moniker", "time", "HighCorrPositive", "HighCorrNegative", "Enthusiastic", "Relaxed", "Sad", "Nervous", "Stressed", "MWoccur", "EmotionChronometry")
    data <- data[corrred]

# Remove linear trends in data
scaled.dat <- scale(data[,3:11]) #standardize
names <- data[,1:2]
scaled.dat <- cbind(names, scaled.dat)
data <- scaled.dat

for(i in levels(data$moniker)) {
    data1 <- data[data$moniker == i,]
                data1$Enthusiastic <- detrend(data1$Enthusiastic, tt = 'linear')
                data1$Relaxed <- detrend(data1$Relaxed, tt = 'linear')
                data1$Nervous <- detrend(data1$Nervous, tt = 'linear')
                data1$Sad <- detrend(data1$Sad, tt = 'linear')
                data1$HighCorrPositive <- detrend(data1$HighCorrPositive, tt = 'linear') #Happy, Pleased collapsed
                data1$HighCorrNegative <- detrend(data1$HighCorrNegative, tt = 'linear') #Angry, Irritated collapsed
                data1$Stressed <- detrend(data1$Stressed, tt = 'linear')
                data1$MWoccur <- detrend(data1$MWoccur, tt = 'linear')
                data1$EmotionChronometry <- detrend(data1$EmotionChronometry, tt = 'linear')
    savefile <- paste0 ("~/step2/", i, ".csv", sep = "") #save next preprocessed step here to a folder you name 'step2' (or change the name to what you would like)
    write.csv(data1, file = savefile, row.names = FALSE)
    print(paste("Dataframe Saved:", i))
}

# Make sure timing is equally spaced, provide the overnight NA value for overnight spacing of self-report and diurnal time, or exogneous time variable taken as square root of time of day.
datadetrend <- rbindlist(lapply(list.files("~/step2/", full.names = TRUE), fread), fill = TRUE) #read in individual files from prior step
    datadetrend$moniker <- as.factor(datadetrend$moniker)
diurnaltime <- read.csv("~/overnightanddiurnaltime.csv") #time=ESM timepoint, contime=the continuous order of timepoint occurrence for ordering, raw time=timepoint by hours in the day, TimeofDay=diurnal time calculated by square root of time based on literature.
diurnaltime <- diurnaltime[,-3]
for(i in levels(datadetrend$moniker)) {
    datad1 <- datadetrend[datadetrend$moniker == i,]
          datad1 <- merge(diurnaltime, datad1, by = "time", all = TRUE)
          datad1$moniker[is.na(datad1$moniker)] <- datad1$moniker[1]
          datad1 <- datad1[order(datad1$contime),]
          remove <- c(1,2,3) #remove original time, continuous time, and moniker in order to feed into GIMME
          datad1 <- datad1[, -(remove)]
    savefile <- paste0 ("~/finaloutput/", i, ".csv", sep = "") #save final preprocessed step here to a folder you name 'finaloutput' (or change the name to what you would like)
    write.csv(datad1, file = savefile, row.names = FALSE)
    print(paste("Dataframe Saved:", i))
}
## Pre-processing Complete
```

#2. Confirmatory groups GIMME analysis - clinical diagnostic subgroups
```{r}
########################################################
#   CONFIRMATORY 2-GROUPS: ALL CLINICAL vs. HEALTHY    #
########################################################
csgimme2group <- read.csv("~/CS-2gimmeGroups.csv", header = FALSE) #read in clinical group assignments
csgimme2group <- as.data.frame(csgimme2group)

outputcs2 <- gimme(data = "~/finaloutput", #folder with individual pre-processed data files named by ID
    out = "~/2CS-GIMME_output",  #folder to save output
    sep = ",",            
    header = TRUE,        
    subgroup = TRUE,        
    confirm_subgroup = csgimme2group, #confirmatory assignments of clinical diagnostic group
    exogenous = "TimeofDay", #diurnal time of day
    groupcutoff = .75,       
    subcutoff = .51         
    )        

########################################################
#   CONFIRMATORY 3-GROUPS: MDD vs. BPD vs. HEALTHY     #
########################################################
csgimme3group <- read.csv("~/CS-3gimmeGroups.csv", header = FALSE)
csgimme3group <- as.data.frame(csgimme3group)

outputcs3 <- gimme(data = "~/finaloutput",
    out = "~/3CS-GIMME_output",  
    sep = ",",            
    header = TRUE,        
    subgroup = TRUE,        
    confirm_subgroup = csgimme3group, 
    exogenous = "TimeofDay",
    groupcutoff = .75,       
    subcutoff = .51         
    )    

###############################################################
#   CONFIRMATORY 4-GROUPS: MDD vs. REM vs. BPD vs. HEALTHY    #
###############################################################
csgimme4group <- read.csv("~/CS-gimme_groupassignments/CS-4gimmeGroups.csv", header = FALSE)
csgimme4group <- as.data.frame(csgimme4group)

outputcs4 <- gimme(data = "~/finaloutput",
    out = "~/4CS-GIMME_output",  
    sep = ",",            
    header = TRUE,        
    subgroup = TRUE,        
    confirm_subgroup = csgimme4group, 
    exogenous = "TimeofDay",
    groupcutoff = .75,       
    subcutoff = .51         
    )    
```

#3. Data-driven groups GIMME analysis - data-driven subgroups
```{r}
##################################
##     S-GIMME, DATA-DRIVEN     ##
##################################

sgimmefit <- gimme(data = "~/finaloutput", #folder with individual pre-processed data files named by ID
         out = "~/datadrivenGIMME_output", #folder to save output
         sep = ",",
         header = TRUE,
         ar = TRUE,
         plot = TRUE,
         subgroup = TRUE,
         paths = NULL, 
         exogenous = "TimeofDay",
         groupcutoff = .75, 
         subcutoff   = .51) 
```

#4. Final GIMME solution evaluation
```{r}
##########
#ROBUSTNESS VALUES FOR S-GIMME SUBGROUP SOLUTION
##########
#############################################################################
#method 1: looking at robustness of solution to minor perturbation of edges
#############################################################################
similarity <- as.matrix(read.csv("~/datadrivenGIMME_output/similarityMatrix.csv", header = F)) #read in similarity matrix
evaluatecorr1 <- perturbR(sym.matrix = similarity)
#The VI value for when 20% of nodes are randomly switched:
evaluatecorr1$vi20mark
#The index for when this occurred for the first time in results
min(which(colMeans(evaluatecorr1$VI)>evaluatecorr1$vi20mark))
#The alpha/percent that corresponds with this index:
evaluatecorr1$percent[min(which(colMeans(evaluatecorr1$VI)>evaluatecorr1$vi20mark))]
#the ARI values:
evaluatecorr1$ari20mark 
mean(evaluatecorr1$ARI[,which(round(evaluatecorr1$percent, digits = 2) == .20)])

#########################################################
#method 2: Look at distribution of modularity values. ##
#########################################################
#The modularity value from the solution on the original matrix:
evaluatecorr1$modularity[1,1]
#against the distribution of randomly generated graphs (matrices) with similar properties to ours:
hist(evaluatecorr1$modularity[,which(round(evaluatecorr1$percent, digits = 2) ==1.00)],
xlim = c(0,1), xlab = "Modularity Values", main = "Histogram of Modularity Values")
abline(v = evaluatecorr1$modularity[1,1], col = "red")
evaluatecorr1$cutoff
```